---
title: 'Analysis and visualisations for "Reproducible research and GIScience: an evaluation using AGILE conference papers"'
author: "Daniel Nüst, Barbara Hofer"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    keep_tex: yes
    latex_engine: xelatex
    toc: yes
  html_document:
    df_print: paged
    toc: yes
urlcolor: blue
# add lscape package to support kableExtra::landscape() for PDF output
header-includes: |
  \usepackage{lscape}
---

## License

This document is licensed under a [Creative Commons Attribution 4.0 International License](https://creativecommons.org/licenses/by/4.0/).

All contained code is licensed under the [Apache License 2.0](https://choosealicense.com/licenses/apache-2.0/).

The data used is licensed under a [Open Data Commons Attribution License](https://opendatacommons.org/licenses/by/).

See the paper's "Author Contributions" section for details on the contributors of data files.

\newpage

## Metadata

Required libraries and runtime environment description.

```{r load_libraries, echo=TRUE, message=FALSE, warning=FALSE}
library("pdftools")
library("stringr")
library("tidyverse")
library("knitr")
library("tidytext")
library("wordcloud")
library("RColorBrewer")
library("readr")
library("ggplot2")
library("rvest")
library("jsonlite")
library("reshape2")
library("ggthemes")
library("grid")
library("gridBase")
library("gridExtra")
library("kableExtra")
library("devtools")
library("rlang")
library("huxtable")
library("here")
library("httr")
```

```{r session_info}
devtools::session_info(include_base = TRUE)
```

This document is versioned in a public [git](https://git-scm.com/) repository, [https://github.com/nuest/reproducible-research-and-giscience](https://github.com/nuest/reproducible-research-and-giscience). The current revision is `r system2("git", "rev-parse --short HEAD", stdout = TRUE, stderr = TRUE)` with the version tag `r system2("git", "tag --list --points-at HEAD", stdout = TRUE, stderr = TRUE)`.

\newpage

## Prerequisites

### API key

An API key is needed for accessing the [Springer API](https://dev.springer.com/) to automatically retrieve the number of full papers.
Create a file `.Renviron` next to this document and add the following line:

```
SPRINGER_API_KEY=<your key>
```

Or set the environment variable within this notebook:

```{r set_api_key,eval=FALSE}
Sys.setenv(SPRINGER_API_KEY = "<your key>")
```

```{r check_api_key}
if (is.na(Sys.getenv("SPRINGER_API_KEY", unset = NA)))
  warning("API key is not set, please check section \"Prerequisites\" of the Rmd file.")
```

```{r data_path}
data_path <- "paper-corpus"
```

### Data

The data for the analysis is required in form of a directory with PDF files.
Add the PDFs to a directory called ` `r data_path` ` (path automatically inserted here based on above variable) next to the file `agile-rr-paper-corpus.Rmd` (this file).

You can contact the original paper authors and ask for the test dataset to reproduce the full analysis.
Alternatively, you can download a selection of AGILE short papers to test the workflow using the code below which is _not_ executed by default.

```{r demo_data,eval=FALSE}
dir.create(here::here(data_path))

# harvest links to PDFs, select more years for more data, 
# e.g. c(2003:2017) and increase max_files_per_year
years <- c(2015:2017)
max_files_per_year <- 10
base_url <- "https://agile-online.org/index.php/conference/proceedings/proceedings-"
proceedings_urls <- sapply(X = as.character(years),
                           FUN = function(x) { paste0(base_url, x) }, USE.NAMES=TRUE)
proceedings_html <- lapply(X = proceedings_urls, FUN = read_html)

# papers, posters, abstracts of full papers, keynotes - we don't care as long it is pdf
# we might also catch both abstract of a poster and the poster itself
get_links <- function(page){ 
  all_links <- page %>%
    html_nodes(css = "a") %>%
    html_attr("href") %>%
    as.list()
  pdf_links <- tibble(links = all_links) %>%
    filter(str_detect(links, pattern = "pdf$"))
  return(pdf_links)
}

proceedings_links_any <- lapply(X = proceedings_html, FUN = get_links)

base_url <- "https://agile-online.org/"
files <- lapply(X = names(proceedings_links_any), FUN = function(x) {
  year <- x
  file_in_year <- 1
  max_files <- min(max_files_per_year, length(proceedings_links_any[[year]]$links))
  year_links <- proceedings_links_any[[year]]$links[c(1:max_files)]
  
  files <- lapply(X= year_links, FUN = function(x) {
    link_url <- paste0(base_url, x)
    filename <- here::here(data_path, 
                           paste0(year, file_in_year, "_", basename(x)))
    if(!file.exists(filename)) {
      response <- GET(url = link_url)
      raw_content <- content(response, "raw")
      writeBin(raw_content, filename)
      #cat("Saved URL", link_url, "\t\tto file\t\t", filename, "\n")
    }
    filename
    file_in_year <<- file_in_year + 1
  })
  files
  cat("Downloaded", length(files), "files for year", year, "\n")
})
```

### Code

The **text analysis** is based the R package [`tidytext`](https://cran.r-project.org/package=tidytext) from the [`tidyverse`](https://www.tidyverse.org/) suite of packages and uses the [`dplyr`](http://dplyr.tidyverse.org/) grammar.
Read the [`tidytext` tutorial](https://cran.r-project.org/web/packages/tidytext/vignettes/tidytext.html) to learn about the used functions and concepts.

The **plots and tables** of survey data and evaluation use the packages [`ggplot2`](http://ggplot2.tidyverse.org/), [`knitr::kable()`](https://yihui.name/knitr/), [`huxtable`](https://hughjonesd.github.io/huxtable/), and [`kableExtra`](https://cran.r-project.org/package=kableExtra).

\newpage

## Reproduce paper

_If you do not have the original data or do not download the data, you cannot reproduce the text analysis part of the paper, i.e. wordcloud and terms frequency analysis._ **You can still reproduce the other figures**.

To create the PDF of the reproducibility package based on this document you can run the following commandsin a new R session after completing the prerequisites with the original paper corpus data.

```{r render_with_rmarkdown,eval=FALSE}
require("knitr")
require("rmarkdown")
rmarkdown::render("agile-rr-paper-corpus.Rmd", output_format = "pdf_document")
```

\newpage

## Paper corpus: loading and cleaning

The test dataset for the analysis cannot be shared publicly due to copyrights.
It comprises all nominees for the best paper award since 2008, both short papers and full papers.
See the paper supplemental files for a full list of citations.

The analysis loads all files from the directory _`r here::here(data_path)`_.

```{r load_filenames}
files <- dir(path = here::here(data_path), pattern = ".pdf$", full.names = TRUE)
```

This analysis was created with the following `r length(files)` documents, `r length(which(grepl("shortpaper", files)))` of which are short papers:

```{r list_files,echo=FALSE}
# remove base name
sapply(X = files, FUN = stringr::str_remove, USE.NAMES = FALSE, pattern = here())
```

Read the data from PDFs and preprocess to create a [tidy](https://www.jstatsoft.org/article/view/v059i10) data structure without [stop words](https://en.wikipedia.org/wiki/Stop_words):

```{r tidy_data}
texts <- lapply(files, pdf_text)
texts <- unlist(lapply(texts, str_c, collapse = TRUE))
infos <- lapply(files, pdf_info)

if (!is.null(texts)) {
  tidy_texts <- tibble(id = str_extract(files, "[0-9]+"),
                       file = files,
                       text = texts,
                       pages = map_chr(infos, function(info) {info$pages}))
  
  papers_words <- tidy_texts %>%
    select(file,
           text) %>%
    unnest_tokens(word, text)
  
  my_stop_words <- tibble(
    word = c(
      "et",
      "al",
      "fig",
      "e.g",
      "i.e",
      "http",
      "ing",
      "pp",
      "figure",
      "based"
    ),
    lexicon = "agile"
  )
  all_stop_words <- stop_words %>%
    bind_rows(my_stop_words)
  
  suppressWarnings({
    no_numbers <- papers_words %>%
      filter(is.na(as.numeric(word)))
  })
  no_stop_words <- no_numbers %>%
    anti_join(all_stop_words, by = "word") %>%
    mutate(id = str_extract(file, "[0-9]+"))
} else {
  warning("No input data provided at ", here::here(data_path))
  # create empty outputs if no input data is given
  papers_words <- tibble(word = c("no data"))
  no_stop_words <- tibble(id = c("no data"), word = c("no data"))
  tidy_texts <- tibble(id = c("no data"))
}
```

```{r calculate_stopword_stats, echo=FALSE}
total_words = nrow(papers_words)
after_cleanup = nrow(no_stop_words)
```

About `r round(after_cleanup/total_words * 100)` % of the words are considered stop words.

\newpage

_How many non-stop words does each document have?_

```{r stop_words}
kable(no_stop_words %>%
  group_by(id) %>%
  summarise(words = n()) %>%
  arrange(desc(words)))
```

**Note:** In the original paper corpus there was an issue with reading in one paper, which only had 15 words. Since it was not possible to copy or extract text, it was send through an OCR process (using [OCRmyPDF](https://github.com/jbarlow83/OCRmyPDF)) with the command `docker run -v $(pwd)/paper-corpus:/home/docker -it jbarlow83/ocrmypdf-tess4 --force-ocr 22015_Mazimpaka_Timpf_AGILE.pdf 22015_Mazimpaka_Timpf_AGILE_ocr.pdf` and the created file was used instead of the original.

\newpage

## Table: Reproducible research-related keywords in the corpus

_How often do the following terms appear in each paper?_

The detection matches full words using regex option `\b`.

- reproduc (``, reproducibility, reproducible, reproduce, reproduction)
- replic (`replicat.*`, i.e. replication, replicate)
- repeatab (`repeatab.*`, i.e. repeatability, repeatable)
- software
- (pseudo) code/script(s) [column name _code_]
- algorithm (`algorithm.*`, i.e. algorithms, algorithmic)
- process (`process.*`, i.e. processing, processes, preprocessing)
- data (`data.*`, i.e. dataset(s), database(s))
- result(s)
- repository(ies)

```{r keywords_per_paper}
tidy_texts_lower <- str_to_lower(tidy_texts$text)
word_counts <- tibble(
  id = tidy_texts$id,
  `reproduc..` = str_count(tidy_texts_lower, "\\breproduc.*\\b"),
  `replic..` = str_count(tidy_texts_lower, "\\breplicat.*\\b"),
  `repeatab..` = str_count(tidy_texts_lower, "\\brepeatab.*\\b"),
  `code` = str_count(tidy_texts_lower,
    "(\\bcode\\b|\\bscript.*\\b|\\bpseudo\ code\\b)"),
  software = str_count(tidy_texts_lower, "\\bsoftware\\b"),
  `algorithm(s)` = str_count(tidy_texts_lower, "\\balgorithm.*\\b"),
  `(pre)process..` = str_count(tidy_texts_lower, 
                "(\\bprocess.*\\b|\\bpreprocess.*\\b|\\bpre-process.*\\b)"),
  `data.*` = str_count(tidy_texts_lower, "\\bdata.*\\b"),
  `result(s)` = str_count(tidy_texts_lower, "\\bresults?\\b"),
  `repository/ies` = str_count(tidy_texts_lower, "\\brepositor(y|ies)\\b")
)

# https://stackoverflow.com/a/32827260/261210
sumColsInARow <- function(df, list_of_cols, new_col) {
  df %>% 
    mutate_(.dots = ~Reduce(`+`, .[list_of_cols])) %>% 
    setNames(c(names(df), new_col))
}

word_counts_sums <- sumColsInARow(
  word_counts, 
  names(word_counts)[names(word_counts) != "id"], "all") %>%
  arrange(desc(all))

# load paper names from evaluation table
citations <- read_csv("Paper_Evaluation.csv", 
                      col_types = cols_only(author = col_character(),
                                            paper = col_character()))
word_counts_sums <- word_counts_sums %>%
  left_join(citations, by = c("id" = "paper")) %>%
  select(citation = author, `reproduc..`:`result(s)`, `all`)

word_counts_sums_total <- word_counts_sums %>% 
  summarise_if(is.numeric, funs(sum)) %>%
  add_column(citation = "Total", .before = 0)
word_counts_sums <- rbind(word_counts_sums, word_counts_sums_total)

# for inline testing: kable(word_counts_sums)
kable(word_counts_sums,
      caption = paste0("Reproducible research-related keywords in the corpus,",
                       " ordered by sum of matches per paper"),
      format = "latex", booktabs = TRUE) %>%
  kableExtra::landscape()
```

\newpage

## Figure: Word cloud of test corpus papers (A), and top words (B)

```{r top_words}
countPapersUsingWord <- function(the_word) {
  sapply(the_word, function(w) {
    no_stop_words %>%
      filter(word == w) %>%
      group_by(id) %>%
      count %>%
      nrow
  })
}

top_words <- no_stop_words %>%
  group_by(word) %>%
  tally %>%
  arrange(desc(n)) %>%
  head(20) %>%
  mutate(`# papers` = countPapersUsingWord(word)) %>%
  add_column(place = c(1:nrow(.)), .before = 0)
```

```{r Fig1,dpi=600,fig.width=7,fig.asp=0.85}
set.seed(1)
if (max(top_words$n) < 100) {
  minimum_occurence <- round(mean(top_words$n))
} else {
  minimum_occurence <- 100
}
cloud_words <- no_stop_words %>%
  group_by(word) %>%
  tally %>%
  filter(n >= minimum_occurence) %>% # 100 chosen manually
  arrange(desc(n))

if (nrow(cloud_words) > 0) {  
  def.par <- par(no.readonly = TRUE)
  par(mar = rep(0,4))
  nf <- layout(mat = matrix(data = c(1,2,3,4), nrow = 2, ncol = 2, byrow = TRUE),
         widths = c(lcm(8),lcm(8)),
         heights = c(lcm(0.5),lcm(11)))
  #layout.show(nf)
  plot.new()
  text(0.5, 0.5, "A", font = 2)
  plot.new()
  text(0.5, 0.5, "B", font = 2)
  
  wordcloud(cloud_words$word, cloud_words$n,
            max.words = Inf,
            random.order = FALSE,
            fixed.asp = FALSE,
            rot.per = 0,
            color = brewer.pal(8,"Dark2"))
  
  frame() # thx to https://stackoverflow.com/a/25194694/261210
  vps <- baseViewports()
  pushViewport(vps$inner, vps$figure, vps$plot)
  grid.table(as.matrix(top_words),
             theme = ttheme_minimal(base_size = 11,
                                    padding = unit(c(10,5), "pt"))
             )
  popViewport(3)

  par(def.par)
} else {
  warning("No input data for wordcloud provided")
}
```

This word cloud is based on `r length(unique(cloud_words$word))` unique words occuring each at least `r minimum_occurence` times, all in all occuring `r sum(cloud_words$n)` times which comprises `r round(sum(cloud_words$n)/ nrow(no_stop_words) * 100)` % of non-stop words.

\newpage

## Reproduciblity assessment

```{r evaldata_file}
evaldata_file <- "Paper_Evaluation.csv"
```

The following plots are based on the file `r evaldata_file`, the result from the manual reproducibility assessment.

```{r load_evaldata,warning=FALSE}
category_levels <- c("0", "1", "2", "3")
paper_evaluation_raw <- read_csv(evaldata_file, 
    col_types = cols(
      paper = col_skip(),
      title = col_skip(),
      `Notes Reviewer` = col_skip(),
      `computational environment` = col_factor(levels = category_levels),
      `input data` = col_factor(levels = category_levels),
      `method/analysis/processing` = col_factor(levels = category_levels),
      preprocessing = col_factor(levels = category_levels),
      results = col_factor(levels = category_levels),
      X12 = col_skip(),
      X14 = col_skip(),
      `Notes Reviewer` = col_skip(),
      `Author comment` = col_skip()
      ),
    na = "NA")
categoryColumns <- c("input data", 
                     "preprocessing",
                     "method/analysis/processing",
                     "computational environment",
                     "results")
```

```{r corpus_table_with_small_font_for_latex}
options(knitr.kable.NA = '-')
kable(paper_evaluation_raw %>% 
        select(-matches("reviewer")) %>%
        mutate(`short paper` = if_else(`short paper` == TRUE, "X", "")),
      format = "latex", booktabs = TRUE,
      caption = paste0("Reproducibility levels for paper corpus; ",
                       "'-' is category not available")) %>%
  kable_styling(latex_options = "scale_down")
```

\newpage

## Conceptual papers

```{r conceptual_papers,warning=FALSE}
paper_evaluation <- paper_evaluation_raw %>%
  # add year column
  mutate(year = as.numeric(str_extract(author, "[0-9]+"))) %>%
  # create new attribute for conceptual papers
  mutate(conceptual = is.na(`input data`) 
         & is.na(preprocessing) 
         & is.na(`method/analysis/processing`) 
         & is.na(`computational environment`) 
         & is.na(results))

count_conceptual <- nrow(paper_evaluation %>% 
                           filter(conceptual))
count_mixed <- nrow(paper_evaluation %>% 
                      filter(is.na(`input data`) 
                             | is.na(preprocessing) 
                             | is.na(`method/analysis/processing`) 
                             | is.na(`computational environment`) 
                             | is.na(results)))
```

`r count_conceptual` papers are purely conceptual (all categories have value `NA`).
These are not included in the following statistics.

`r count_mixed` papers are partically conceptual (at least one category has a value of `NA`).
These are evaluated.

`r paper_evaluation %>% filter(is.na(preprocessing)) %>% count() %>% .$n` papers are not applicable for preprocessing criterion.

\newpage

## Overall conference contributions

_How many conference contributions were made at AGILE conferences over the years?_

We need to scrape data from the AGILE website for short papers and posters.

```{r harvest_agile_website, cache=TRUE}
base_url <- "https://agile-online.org/index.php/conference/proceedings/proceedings-"
proceedings_urls <- sapply(X = as.character(c(2003:2017)), 
                           FUN = function(x) { paste0(base_url, x)}, 
                           USE.NAMES = TRUE)
proceedings_html <- lapply(X = proceedings_urls, FUN = read_html)

get_paper_links <- function(page){ 
  links <- page %>%
    html_nodes(css = "a") %>%
    html_attr("href") %>%
    as.list() %>%
    tibble(links = .) %>%
    filter(str_detect(links, 
                      pattern = "(ShortPapers|papers|proceedings|papers/Paper_)/[^pP]"))
  return(links)
}

# papers, posters, abstracts of full papers - we don't care as long it is pdf
get_all_links <- function(page){ 
  all_links <- page %>%
    html_nodes(css = "a") %>%
    html_attr("href") %>%
    as.list()
  
  pdf_links <- tibble(links = all_links) %>%
    filter(str_detect(links, pattern = "pdf$")) %>%
    # keep only one of poster abstract and poster PDF:
    filter(!str_detect(links, pattern = "Poster_in_PDF.pdf")) %>%
    # some keynotes are also available for Download (at least one in 2012), remove them:
    filter(!str_detect(links, pattern = "(keynotes|Keynote)"))
  
  return(pdf_links)
}

get_non_full_papers_links <- function(page){ 
  get_all_links(page) %>%
    # 2017 includes full paper abstracts in the PDFs, remove them:
    filter(!str_detect(links, pattern = "FullPaperAbstract"))
}

proceedings_links_short_and_full_papers <- lapply(X = proceedings_html, 
                                                  FUN = get_non_full_papers_links)
```

Get the ISBNs of AGILE proceedings via harvesting AGILE and Springer websites.
Then query [Springer API](https://dev.springer.com/) for number of chapters in each book to get the full paper count.

```{r harvest_springer_api,cache=TRUE}
if(is.na(Sys.getenv("SPRINGER_API_KEY", unset = NA))) {
  # no API key provided, add some dummy data for the document to render
  all_contributions <- NA
  full_papers <- NA
  paper_counts <- tibble(year = c(NA))
  sample_full_papers <- NA
  sample_short_papers <- NA
} else {
  base_url_lngc <- "https://agile-online.org/index.php/conference/springer-series"
  # 2007 and 2017 are missing on the AGILE website
  lngc_2007 <- "https://link.springer.com/book/10.1007%2F978-3-540-72385-1"
  lngc_2017 <- "https://link.springer.com/book/10.1007/978-3-319-56759-4"
  
  springer_api_key <- paste0("&api_key=", Sys.getenv("SPRINGER_API_KEY"))
  springer_api_base <- "http://api.springer.com/metadata/json?"
  
  lngc_html <- read_html(base_url_lngc)
  
  lngc_books_urls <- lngc_html %>%
    html_nodes(css = "a") %>%
    html_attr("href") %>%
    tibble(links = .) %>%
    filter(str_detect(links, pattern = "/book/")) %>%
    add_row(links = lngc_2007) %>%
    add_row(links = lngc_2017)
  
  get_full_paper_count <- function(link) {
    # extract id for book
    isbn <- read_html(link) %>%
      html_nodes("span[id=print-isbn], dd[itemprop=isbn]") %>%
      html_text()
    year <- read_html(link) %>%
      html_nodes("span[id=copyright-info], div[class=copyright]") %>%
      html_text() %>%
      gsub("[^0-9]", "", .) %>%
      as.numeric(.)
    
    url <- str_c(springer_api_base, "q=isbn:", isbn, springer_api_key)
    
    #cat("Query with isbn ", isbn, " for year ", year, ": ", url, "... ")
    metadata <- fromJSON(url)
    total <- as.numeric(metadata$result$total)
    #cat("Result: ", total, "\n")
    return(tibble(year = year, `full paper` = total))
  }
  
  lngc_full_paper_counts <- bind_rows(lapply(lngc_books_urls$links, get_full_paper_count))
  
  counts_any <- sapply(proceedings_links_short_and_full_papers, 
                       function(x) { length(x[["links"]]) })
  non_full_paper_counts <- tibble(
    year = as.numeric(names(counts_any)), 
    `short paper/poster` = counts_any)
  
  paper_counts <- full_join(lngc_full_paper_counts, non_full_paper_counts, by = "year") %>%
    arrange(desc(year))
  
  all_contributions <- 
    sum(paper_counts$"full paper", na.rm = TRUE) + 
    sum(paper_counts$"short paper/poster", na.rm = TRUE)
  full_papers <- sum(paper_counts$"full paper", na.rm = TRUE)
  
  sample_full_papers <- paper_evaluation %>% 
    filter(`short paper` == FALSE) %>% 
    count() %>% 
    .$n
  sample_short_papers <- paper_evaluation %>% 
    filter(`short paper` == TRUE) %>% 
    count() %>% 
    .$n
  
  kable(paper_counts)
}
```

Overall **`r all_contributions` conference contributions** (including posters and short papers), of which **`r full_papers` are full papers**, in the years `r min(paper_counts$year)` to `r max(paper_counts$year)`.

The used **sample** contains `r sample_full_papers` full papers (`r round(sample_full_papers / full_papers * 100, digits = 2)` %) and `r sample_short_papers` short papers (percentage respectively full number of short papers not available because not distinguishable from poster abstracts for some years).

\newpage

## Table: Statistics of reproducibility levels per criterion

```{r summary_evaldata}
evaldata_numeric <- paper_evaluation %>%
  # must convert factors to numbers to calculate the mean and median
  mutate_if(is.factor, funs(as.integer(as.character(.))))

summary(evaldata_numeric[,categoryColumns])

# apply summary independently to format as table
summaries <- sapply(evaldata_numeric[,categoryColumns], summary)
exclude_values_summary <- c("1st Qu.", "3rd Qu.")
kable(subset(summaries, !(rownames(summaries) %in% exclude_values_summary)), 
      digits = 2,
      col.names = c("input data", "preproc.", "method/analysis/proc.",
                    "comp. env.", "results"),
      caption = paste0("\\label{tab:levels_statistics}Statistics of ",
                       "reproducibility levels per criterion"))
```

The preprocessing has `r sum(!is.na(evaldata_numeric$preprocessing))` values, with `0` and `1` around the "middle" resulting in a fraction as the median.

\newpage

## Figure: Results of reproducibility assessment

```{r Fig3,fig.width=10}
# match the colours to time series plot below
colours <- RColorBrewer::brewer.pal(length(categoryColumns), "Set1")
level_names <- c("0", "1", "2", "3", "NA")

criteriaBarplot = function(data, main, colour) {
  barplot(table(data, useNA = "always"), 
          main = main,
          xlab = "Level", 
          ylim = c(0,25),
          names.arg = level_names,col = colours[colour])
}

par(mfrow = c(1,length(categoryColumns)))
criteriaBarplot(paper_evaluation$`input data`,
                main = "A: Input data", colour = 1)
criteriaBarplot(paper_evaluation$`preprocessing`, 
                main = "B: Preprocessing", colour = 2)
criteriaBarplot(paper_evaluation$`method/analysis/processing`,
                main = "C: Methods/Analysis/\nProcessing", colour = 3)
criteriaBarplot(paper_evaluation$`computational environment`,
                main = "D: Computational\nEnvironment", colour = 4)
criteriaBarplot(paper_evaluation$results,
                main = "E: Results", colour = 5)
```

```{r criteria_numbers}
data_level_zero <- paper_evaluation %>% 
  filter(`input data` == 0) %>% 
  count() %>% .$n

data_level_two <- paper_evaluation %>% 
  filter(`input data` == 2) %>% 
  count() %>% .$n

preprocessing_included <- paper_evaluation %>% 
  filter(!is.na(preprocessing)) %>% 
  count() %>% .$n

methods_and_results_eq_one <- evaldata_numeric %>% 
  filter(`method/analysis/processing` == 1 & results == 1) %>% 
  count() %>% .$n
```

`r data_level_zero` papers have level `0` and `r data_level_two` have level `2` in the data criterion.

`r preprocessing_included` papers include some kind of preprocessing.

`r methods_and_results_eq_one` papers have level `1` in both methods and results criterion.

\newpage

## Table: Mean levels per criterion for full and short papers

```{r summary_evaldata_grouped}
summaries_short_paper <- sapply(evaldata_numeric %>%
                                  filter(`short paper` == TRUE) %>%
                                  select(categoryColumns), summary)
means_short_paper <- subset(summaries_short_paper, rownames(summaries) %in% c("Mean"))
rownames(means_short_paper) <- c("Short papers")
summaries_full_paper <- sapply(evaldata_numeric %>% filter(`short paper` == FALSE) %>%
                                 select(categoryColumns), summary)
means_full_paper <- subset(summaries_full_paper, rownames(summaries) %in% c("Mean"))
rownames(means_full_paper) <- c("Full papers")
```

\small

```{r summary_evaldata_grouped_smallfont_latex}
kable(rbind(means_full_paper, means_short_paper),
      digits = 2,
      col.names = c("input data", "preproc.", "method/analysis/proc.", "comp. env.", "results"),
      caption = paste0("\\label{tab:mean_full_vs_short}",
                       "Mean levels per criterion for full and short papers"))
```

\normalsize

\newpage

## Extra table: Mean levels averaged across criteria over time

```{r evaldata_summary_by_year_mean}
means_years <- evaldata_numeric %>%
  filter(conceptual == FALSE) %>%
  group_by(year) %>%
  summarise(mean = mean(c(`input data`, 
                          preprocessing, 
                          `method/analysis/processing`, 
                          `computational environment`, 
                          `results`),
                        na.rm = TRUE),
            `paper count` = n())

means_years_table <- means_years %>% 
        mutate(mean = round(mean, 2), 
               `paper count` = as.character(`paper count`)) %>%
        mutate(labels = str_c(year, " (n = ", `paper count`, ")")) %>%
        #column_to_rownames("labels") %>%
        select(mean) %>%
        t()
```

\small

```{r summary_by_year_smallfont_latex}
kable(means_years_table,
      caption = "Summarised mean values over all criteria over time")
```

\normalsize

\newpage

## Figure: Mean reproducibility levels per category over time

```{r Fig4,fig.width=10,dpi=300}
evaldata_years <- evaldata_numeric %>%
  filter(conceptual == FALSE) %>%
  filter(year != 2011) %>%
  group_by(year) %>%
  summarise(input = mean(`input data`, na.rm = TRUE),
         preprocessing = mean(preprocessing, na.rm = TRUE),
            method = mean(`method/analysis/processing`, na.rm = TRUE),
            environment = mean(`computational environment`, na.rm = TRUE),
            results = mean(results, na.rm = TRUE))
paper_count_years <- evaldata_numeric %>%
  filter(conceptual == FALSE) %>%
  filter(year != 2011) %>%
  group_by(year) %>%
  summarise(`paper count` = n())

evaldata_years_long <- melt(evaldata_years, id.vars = c("year"))
ggplot(evaldata_years_long, aes(year, value)) +
  geom_bar(aes(fill = variable), position = "dodge", stat = "identity") +
  ylab("mean value of criterion level") + 
  scale_x_continuous(breaks = evaldata_years$year,
                     labels = paste0(paper_count_years$year, 
                                     " (n=", 
                                     paper_count_years$`paper count`, 
                                     ")")) +
  scale_fill_brewer(palette = "Set1", name = "Category") +
  theme_tufte(base_size = 18) +
  theme(legend.position = c(0.15,0.75), 
        legend.text = element_text(size = 14)) +
  ylim(0, 3) +
  stat_summary(fun.y = mean, fun.ymin = mean, fun.ymax = mean, shape = "-", size = 2) +
  stat_summary(fun.y = mean, geom = "line", linetype = "dotted", mapping = aes(group = 1))
```

\newpage

## Figure: Author survey results on the importance of reproducibility

```{r Fig5,warning=FALSE,fig.width=12,dpi=300}
Reproducibility_Survey <- read_delim(file = "Reproducibility_Survey.csv", 
    delim = ";", 
    escape_double = FALSE, 
    col_types = cols(`Short/Full Paper` = col_factor(levels = c("Full", 
        "Short")), 
        Timestamp = col_datetime(format = "%m/%d/%Y %H:%M:%S"), 
        X15 = col_skip()), 
    trim_ws = TRUE) %>%
  rename(`considered reproducibility` =
 `Have you considered the reproducibility of research published in your nominated paper?`)

considered_reproducibility <- Reproducibility_Survey %>%
  group_by(`Short/Full Paper`,
           `considered reproducibility`) %>%
  filter(!is.na(`considered reproducibility`)) %>%
  count()

responses_full <- considered_reproducibility %>% 
  filter(`Short/Full Paper` == "Full") %>% 
  .$n %>% sum()
responses_short <- considered_reproducibility %>% 
  filter(`Short/Full Paper` == "Short") %>% 
  .$n %>% sum()

responses_for_papers_count <- length(
  # substract 1 for "The author has not agreed"
  unique(Reproducibility_Survey$`Please select your nominated AGILE Best Paper.`)) - 1

anonymous_responses_count <- Reproducibility_Survey %>%
  filter(is.na(`considered reproducibility`)) %>%
  count()

ggplot(data = Reproducibility_Survey %>%
  filter(!is.na(`considered reproducibility`)), 
  aes(x = `considered reproducibility`, 
           fill = `Short/Full Paper`)) + 
  geom_bar(width = 0.6, position = "dodge") +
  scale_fill_brewer(palette = "Set1", name = "Publication type") +
  scale_x_discrete(label = function(x) str_wrap(x, width = 20),
                   name = paste0("Have you considered the reproducibility of ",
                                 "research published in your nominated paper? (n = ",
                                 sum(considered_reproducibility$n), ")")) +
  scale_y_discrete(name = "Count", limits = c(0:12)) +
  theme_tufte(base_size = 18) + 
  theme(legend.position = c(0.2,0.8), 
        legend.text = element_text(size = 16), 
        legend.key.size = unit(1, "cm")) +
  geom_hline(yintercept = seq(1:10), col = "white", lwd = 0.5)
```

Of the `r sum(considered_reproducibility$n)` responses the plot is based on, `r responses_short` are short and `r responses_full` full papers.

The `r nrow(Reproducibility_Survey)` responses cover `r responses_for_papers_count` papers and include `r anonymous_responses_count` responses without consent to use the data.

\newpage

## Table: Hindering circumstances for reproducibility for each survey response

```{r survey_results_hindering_circumstances}
hindering_circumstances <- Reproducibility_Survey %>%
  select(starts_with('Please rate')) %>%
  drop_na() %>% # remove responses with no answers
  # order the levels of the factors:
  mutate_all(factor, levels = c("Not at all",
                                "Slightly hindered",
                                "Moderately hindered",
                                "Strongly hindered",
                                "Main reason"), ordered = TRUE)

names(hindering_circumstances) <- sapply(names(hindering_circumstances), function(name) {
  if (grepl(".*legal.*", name, ignore.case = TRUE))
    return("Legal restrictions")
  else if (grepl(pattern = ".*time.*", x = name, ignore.case = TRUE))
    return("Lack of time")
  else if (grepl(pattern = ".*tools.*", x = name, ignore.case = TRUE))
    return("Lack of tools")
  else if (grepl(pattern = ".*motivation*", x = name, ignore.case = TRUE))
    return("Lack of incentive")
  else if (grepl(pattern = ".knowledge.*", x = name, ignore.case = TRUE))
    return("Lack of knowledge")
  else return(NA)
})

# count the occurences of "main reason" for each question
hindering_circumstances %>%
  summarise_all(funs(sum(grepl(pattern = "Main reason", x = .))))
  
main_reason_counts <- as.data.frame(t(hindering_circumstances %>%
          summarise_all(
            funs(sum(grepl(pattern = "Main reason", x = .)))))) %>%
  rename(count = V1) %>%
  rownames_to_column(var = "circumstance") %>%
  arrange(desc(count))

# sort the columns (circumstances) by the number of "main reason" answers
hindering_circumstances <- hindering_circumstances %>%
  select(main_reason_counts$circumstance) %>%
  # sort the rows by the colum with most "main reason" answers
  arrange(desc(!! rlang::sym(main_reason_counts$circumstance[[1]])))

crcmstncs_ht <- huxtable::as_hux(hindering_circumstances)
# configure font size and cell padding
font_size(crcmstncs_ht) <- 8

bg_colors <- brewer.pal(n = 5, name = "GnBu")
crcmstncs_ht <- crcmstncs_ht %>%
  # set background colors for cells
  set_background_color(where(crcmstncs_ht == "Main reason"), bg_colors[[5]]) %>% 
  set_background_color(where(crcmstncs_ht == "Strongly hindered"), bg_colors[[4]]) %>%
  set_background_color(where(crcmstncs_ht == "Moderately hindered"), bg_colors[[3]]) %>%
  set_background_color(where(crcmstncs_ht == "Slightly hindered"), bg_colors[[2]]) %>%
  set_background_color(where(crcmstncs_ht == "Not at all"), bg_colors[[1]]) %>%
  add_colnames() %>%
  # format column names:
  set_bold(row = 1, col = 1:length(crcmstncs_ht), TRUE) %>%
  set_bottom_border(row = 1, col = 1:length(crcmstncs_ht), 1) %>%
  set_font_size(row = 1, col = 1:length(crcmstncs_ht), value = 10) %>%
  # add label, caption, and float:
  set_label("tab:hindering_circumstances") %>%
  set_latex_float("ht") %>%
  set_width(1) %>%
  set_caption(paste0(
    "Hindering circumstances for reproducibility for each survey response ",
    #"with columns sorted by the respective count of 'main reason' ",
    #"and rows sorted by the answer categories in descending order"
    "(n = ", nrow(hindering_circumstances),
    "); background colour corresponds to cell text."))

crcmstncs_ht
```

