---
title: 'Text-analysis and visualisations for AGILE conference papers 2018-2019'
author: "Daniel Nüst"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output:
  html_document:
    df_print: paged
    toc: yes
abstract: Text analysis of [AGILE conference](https://agile-online.org/conference/) papers - source code at [https://github.com/nuest/reproducible-research-and-giscience](https://github.com/nuest/reproducible-research-and-giscience).
---

## Prerequisites

### Software dependencies

This document does not install the required R packages by default.
You can run the script `install.R` to install all required dependencies on a new R installation, or use `install.packages(..)` to install missing R packages.

```{r install_r, eval=FALSE}
source("install.R")
```

The text analysis is based the R package [`tidytext`](https://cran.r-project.org/package=tidytext) from the [`tidyverse`](https://www.tidyverse.org/) suite of packages and uses the [`dplyr`](http://dplyr.tidyverse.org/) grammar.
Read the [`tidytext` tutorial](https://cran.r-project.org/web/packages/tidytext/vignettes/tidytext.html) to learn about the used functions and concepts.

The plots and tables of survey data and evaluation use the packages [`ggplot2`](http://ggplot2.tidyverse.org/).

Required libraries and runtime environment description.

```{r load_libraries, echo=TRUE, message=FALSE, warning=FALSE}
library("pdftools")
library("stringr")
library("knitr")
library("tibble")
library("tidytext")
library("purrr")
library("dplyr")
library("wordcloud")
library("RColorBrewer")
library("readr")
library("ggplot2")
library("rvest")
library("ggthemes")
library("grid")
library("gridBase")
library("gridExtra")
library("devtools")
library("rlang")
library("huxtable")
library("here")
library("httr")
library("googledrive")
library("SnowballC")
```

### Seed

Seed is set for making word cloud generation reproducible.

```{r seed}
set.seed(1)
```

### Data

```{r data_path}
data_path <- "all-manuscripts"
```

The data for the analysis is required in form of directories with PDF files of all conference papers and poster abstracts.
Due to copyright of full papers, the full paper PDFs must be manually added to the respective directory.
Short papers and poster abstracts are dowloaded automatically.

Add the PDFs to a directory called ` `r data_path` ` this file with one subdirectoy per year:

```{r years}
list.files("all-manuscripts/")
```

The following downloads of AGILE short papers are _not_ executed by default.

```{r data_2018,eval=FALSE}
dir.create(here::here(data_path, "2018"))

page <- read_html("https://agile-online.org/programme-2018/accepted-papers-and-posters-2018")

all_links <- page %>%
    html_nodes(css = "a") %>%
    html_attr("href") %>%
    as.list()

drive_links <- all_links[str_detect(string = all_links, pattern = "drive.google")]
drive_links[sapply(drive_links, is.null)] <- NULL

drive_ids <- lapply(drive_links, as_id)
lapply(drive_ids, drive_download, overwrite = TRUE)
```

```{r data_2019,eval=FALSE}
dir.create(here::here(data_path, "2019"))

page <- read_html("https://agile-online.org/conference-2019/programme-2019/accepted-papers-and-posters-2019")

all_links <- page %>%
    html_nodes(css = "a") %>%
    html_attr("href") %>%
    as.list()

pdf_links <- all_links[str_detect(string = all_links, pattern = ".*Upload_your_PDF.*")]
pdf_links[sapply(pdf_links, is.null)] <- NULL
pdf_links <- paste0("https://agile-online.org", pdf_links)

for (link in pdf_links) {
  download.file(url = link,
                destfile = here::here(data_path,
                                      "2019",
                                      stringr::str_extract(link, "([^/]+$)")))
}
```

## 2018

### Loading and cleaning

```{r load_filenames_2018}
files_2018 <- dir(path = here::here(data_path, "2018"), pattern = ".pdf$", full.names = TRUE)
```

This analysis was created with the following `r length(files_2018)` documents:

```{r list_files_2018,echo=FALSE}
# remove base name
sapply(X = files_2018, FUN = stringr::str_remove, USE.NAMES = FALSE, pattern = here::here(data_path, "2018"))
```

Count the types of submissions:

```{r manuscript_counts_2018}
full_papers_2018 <- length(str_match(files_2018, "10.100")[!is.na(str_match(files_2018, "10.100"))])
other_papers_2018 <- length(str_match(files_2018, "10.100")[is.na(str_match(files_2018, "10.100"))])
```

There are `r full_papers_2018` full papers and `r other_papers_2018` short papers/posters.

Read the data from PDFs and preprocess to create a [tidy](https://www.jstatsoft.org/article/view/v059i10) data structure without [stop words](https://en.wikipedia.org/wiki/Stop_words):

```{r stop_words}
my_stop_words <- tibble(
  word = c(
    "et",
    "al",
    "fig",
    "e.g",
    "i.e",
    "http",
    "ing",
    "pp",
    "figure",
    "table",
    "based",
    "lund", # location of conference 2018
    "https"
  ),
  lexicon = "agile"
)
all_stop_words <- stop_words %>%
  bind_rows(my_stop_words)
```

```{r tidy_data_2018}
texts <- lapply(files_2018, pdf_text)
texts <- unlist(lapply(texts, str_c, collapse = TRUE))
infos <- lapply(files_2018, pdf_info)

make_id <- function(files) {
  str_extract(files, "([^/]+$)")
}

tidy_texts_2018 <- tibble(id = make_id(files_2018),
                     file = files_2018,
                     text = texts,
                     pages = map_chr(infos, function(info) {info$pages}))

papers_words <- tidy_texts_2018 %>%
  select(file, text) %>%
  unnest_tokens(word, text)

suppressWarnings({
  no_numbers <- papers_words %>%
    filter(is.na(as.numeric(word)))
})

no_stop_words_2018 <- no_numbers %>%
  anti_join(all_stop_words, by = "word") %>%
  mutate(id = make_id(file))

# https://github.com/juliasilge/tidytext/issues/17
no_stop_stems_2018 <- no_stop_words_2018 %>%
  mutate(word_stem = wordStem(word))
```

About `r round(nrow(no_stop_words_2018)/nrow(papers_words) * 100)` % of the words are considered stop words.
There are `r length(unique(no_stop_stems_2018$word_stem))` unique word stems of `r length(unique(no_stop_words_2018$word))` words.

_How many non-stop words does each document have?_

```{r stop_words_2018}
no_stop_words_2018 %>%
  group_by(id) %>%
  summarise(words = n()) %>%
  arrange(desc(words))
```

### Text analysis

_How often do the following terms on reproducible research appear in each paper?_

The detection matches full words using regex option `\b`.

- reproduc (`reproduc.*`, reproducibility, reproducible, reproduce, reproduction)
- replic (`replicat.*`, i.e. replication, replicate)
- repeatab (`repeatab.*`, i.e. repeatability, repeatable)
- software
- (pseudo) code/script(s) [column name _code_]
- algorithm (`algorithm.*`, i.e. algorithms, algorithmic)
- process (`process.*`, i.e. processing, processes, preprocessing)
- data (`data.*`, i.e. dataset(s), database(s))
- result(s)
- repository(ies)

```{r keywords_per_paper_2018}
tidy_texts_2018_lower <- str_to_lower(tidy_texts_2018$text)
word_counts <- tibble(
  id = tidy_texts_2018$id,
  `reproduc..` = str_count(tidy_texts_2018_lower, "\\breproduc.*\\b"),
  `replic..` = str_count(tidy_texts_2018_lower, "\\breplicat.*\\b"),
  `repeatab..` = str_count(tidy_texts_2018_lower, "\\brepeatab.*\\b"),
  `code` = str_count(tidy_texts_2018_lower, "(\\bcode\\b|\\bscript.*\\b|\\bpseudo\ code\\b)"),
  `software` = str_count(tidy_texts_2018_lower, "\\bsoftware\\b"),
  `algorithm(s)` = str_count(tidy_texts_2018_lower, "\\balgorithm.*\\b"),
  `(pre)process..` = str_count(tidy_texts_2018_lower, "(\\bprocess.*\\b|\\bpreprocess.*\\b|\\bpre-process.*\\b)"),
  `data.*` = str_count(tidy_texts_2018_lower, "\\bdata.*\\b"),
  `result(s)` = str_count(tidy_texts_2018_lower, "\\bresults?\\b"),
  `repository/ies` = str_count(tidy_texts_2018_lower, "\\brepositor(y|ies)\\b")
) %>%
  mutate(all = rowSums(.[-1]))

word_counts_sums_total_2018 <- word_counts %>% 
  summarise_if(is.numeric, funs(sum)) %>%
  add_column(id = "Total", .before = 0)
rbind(word_counts, word_counts_sums_total_2018)
```

_What are top used words (not stems)?_

```{r top_words_2018}
countPapersUsingWord <- function(the_word) {
  sapply(the_word, function(w) {
    no_stop_words_2018 %>%
      filter(word == w) %>%
      group_by(id) %>%
      count %>%
      nrow
  })
}

top_words_2018 <- no_stop_words_2018 %>%
  group_by(word) %>%
  tally %>%
  arrange(desc(n)) %>%
  head(20) %>%
  mutate(`# papers` = countPapersUsingWord(word)) %>%
  add_column(place = c(1:nrow(.)), .before = 0)

top_words_2018
```

_What are the top word stems?_

```{r top_stems_2018}
countPapersUsingStem <- function(the_stem) {
  sapply(the_stem, function(s) {
    no_stop_stems_2018 %>%
      filter(word_stem == s) %>%
      group_by(id) %>%
      count %>%
      nrow
  })
}

top_stems_2018 <- no_stop_stems_2018 %>%
  group_by(word_stem) %>%
  tally %>%
  arrange(desc(n)) %>%
  head(20) %>%
  mutate(`# papers` = countPapersUsingStem(word_stem)) %>%
  add_column(place = c(1:nrow(.)), .before = 0)

top_stems_2018
```

### Word cloud based on word stems

```{r plot_function}
wordStemPlot <- function(word_stem_data, top_stem_data, year, minimum_occurence, fp_count, op_count) {

  cloud_words <- word_stem_data %>%
    group_by(word_stem) %>%
    tally %>%
    filter(n >= minimum_occurence) %>%
    arrange(desc(n))
  
  def.par <- par(no.readonly = TRUE)
  par(mar = rep(0,4))
  layout(mat = matrix(data = c(1,2,3,4), nrow = 2, ncol = 2, byrow = TRUE),
         widths = c(lcm(8),lcm(8)),
         heights = c(lcm(2),lcm(11)))
  #       -> nf
  #layout.show(nf)
  
  plot.new()
  text(0.5, 0.5, paste0("Word stem cloud of AGILE ", year, " Submissions"), font = 2)
  text(0.5, 0.15, paste0("Based on ", fp_count, " full papers and ", op_count, " short papers/posters.\n",
                        "Showing ", nrow(cloud_words), " of ", sum(cloud_words$n),
                        " word stems occuring at least ", minimum_occurence, " times."), font = 1, cex = 0.7)
  plot.new()
  text(0.5, 0.5, paste0("Top word stems of AGILE ", year, " Submissions"), font = 2)
  text(0.5, 0.15, paste0("Code available at https://github.com/nuest/\nreproducible-research-and-giscience"), font = 1, cex = 0.7)
  
  wordcloud(cloud_words$word_stem, cloud_words$n,
            max.words = Inf,
            random.order = FALSE,
            fixed.asp = FALSE,
            rot.per = 0,
            color = brewer.pal(8,"Dark2"))
  
  frame() # thx to https://stackoverflow.com/a/25194694/261210
  vps <- baseViewports()
  pushViewport(vps$inner, vps$figure, vps$plot)
  grid.table(as.matrix(top_stem_data),
             theme = ttheme_minimal(base_size = 11,
                                    padding = unit(c(10,5), "pt"))
             )
  popViewport(3)
  par(def.par)
}
```

```{r plot_2018,dpi=600,fig.width=7,fig.asp=0.85}
# minimum occurence manually tested so that all words could be plotted
wordStemPlot(no_stop_stems_2018, top_stems_2018, "2018", 200, full_papers_2018, other_papers_2018)
```

--------

## 2019

### Loading and cleaning

```{r load_filenames_2019}
files_2019 <- dir(path = here::here(data_path, "2019"), pattern = ".pdf$", full.names = TRUE)
```

This analysis was created with the following `r length(files_2019)` documents:

```{r list_files_2019,echo=FALSE}
# remove base name
sapply(X = files_2019, FUN = stringr::str_remove, USE.NAMES = FALSE, pattern = here::here(data_path, "2019"))
```

Count the types of submissions:

```{r manuscript_counts_2019}
full_papers_2019 <- length(str_match(files_2019, "10.100")[!is.na(str_match(files_2019, "10.100"))])
other_papers_2019 <- length(str_match(files_2019, "10.100")[is.na(str_match(files_2019, "10.100"))])
```

Read the data from PDFs and preprocess to create a [tidy](https://www.jstatsoft.org/article/view/v059i10) data structure without [stop words](https://en.wikipedia.org/wiki/Stop_words):

```{r tidy_data_2019}
texts <- lapply(files_2019, pdf_text)
texts <- unlist(lapply(texts, str_c, collapse = TRUE))
infos <- lapply(files_2019, pdf_info)

tidy_texts_2019 <- tibble(id = make_id(files_2019),
                     file = files_2019,
                     text = texts,
                     pages = map_chr(infos, function(info) {info$pages}))

papers_words <- tidy_texts_2019 %>%
  select(file, text) %>%
  unnest_tokens(word, text)

suppressWarnings({
  no_numbers <- papers_words %>%
    filter(is.na(as.numeric(word)))
})

no_stop_words_2019 <- no_numbers %>%
  anti_join(all_stop_words, by = "word") %>%
  mutate(id = make_id(file))

# https://github.com/juliasilge/tidytext/issues/17
no_stop_stems_2019 <- no_stop_words_2019 %>%
  mutate(word_stem = wordStem(word))
```

About `r round(nrow(no_stop_words_2019)/nrow(papers_words) * 100)` % of the words are considered stop words.
There are `r length(unique(no_stop_stems_2019$word_stem))` unique word stems of `r length(unique(no_stop_words_2019$word))` words.

**Note:** In the original paper corpus there was an issue with reading in one paper, which only had 1 word, `10.1007@978-3-030-14745-710.pdf`.
Since it was not possible to copy or extract text, it was send through an OCR process (using [OCRmyPDF](https://github.com/jbarlow83/OCRmyPDF)) and the original file renamed to `10.1007@978-3-030-14745-710__pdf`:

```{bash ocr_2019,eval=FALSE}
docker run -v $(pwd)/all-manuscripts/2019:/home/docker -it jbarlow83/ocrmypdf --force-ocr 10.1007@978-3-030-14745-710.pdf 10.1007@978-3-030-14745-710_ocr.pdf
mv all-manuscripts/2019/10.1007@978-3-030-14745-710.pdf all-manuscripts/2019/10.1007@978-3-030-14745-710pdf.orig
```

_How many non-stop words does each document have?_

```{r stop_words_2019}
no_stop_words_2019 %>%
  group_by(id) %>%
  summarise(words = n()) %>%
  arrange(desc(words))
```

### Text analysis

_How often do the following terms on reproducible research appear in each paper?_

The detection matches full words using regex option `\b`.

- reproduc (`reproduc.*`, reproducibility, reproducible, reproduce, reproduction)
- replic (`replicat.*`, i.e. replication, replicate)
- repeatab (`repeatab.*`, i.e. repeatability, repeatable)
- software
- (pseudo) code/script(s) [column name _code_]
- algorithm (`algorithm.*`, i.e. algorithms, algorithmic)
- process (`process.*`, i.e. processing, processes, preprocessing)
- data (`data.*`, i.e. dataset(s), database(s))
- result(s)
- repository(ies)

```{r keywords_per_paper_2019}
tidy_texts_2019_lower <- str_to_lower(tidy_texts_2019$text)
word_counts <- tibble(
  id = tidy_texts_2019$id,
  `reproduc..` = str_count(tidy_texts_2019_lower, "\\breproduc.*\\b"),
  `replic..` = str_count(tidy_texts_2019_lower, "\\breplicat.*\\b"),
  `repeatab..` = str_count(tidy_texts_2019_lower, "\\brepeatab.*\\b"),
  `code` = str_count(tidy_texts_2019_lower, "(\\bcode\\b|\\bscript.*\\b|\\bpseudo\ code\\b)"),
  `software` = str_count(tidy_texts_2019_lower, "\\bsoftware\\b"),
  `algorithm(s)` = str_count(tidy_texts_2019_lower, "\\balgorithm.*\\b"),
  `(pre)process..` = str_count(tidy_texts_2019_lower, "(\\bprocess.*\\b|\\bpreprocess.*\\b|\\bpre-process.*\\b)"),
  `data.*` = str_count(tidy_texts_2019_lower, "\\bdata.*\\b"),
  `result(s)` = str_count(tidy_texts_2019_lower, "\\bresults?\\b"),
  `repository/ies` = str_count(tidy_texts_2019_lower, "\\brepositor(y|ies)\\b")
) %>%
  mutate(all = rowSums(.[-1]))

word_counts_sums_total_2019 <- word_counts %>% 
  summarise_if(is.numeric, funs(sum)) %>%
  add_column(id = "Total", .before = 0)

rbind(word_counts, word_counts_sums_total_2019)
```

_What are top used words (not stems)?_

```{r top_words_2019}
countPapersUsingWord <- function(the_word) {
  sapply(the_word, function(w) {
    no_stop_words_2019 %>%
      filter(word == w) %>%
      group_by(id) %>%
      count %>%
      nrow
  })
}

top_words_2019 <- no_stop_words_2019 %>%
  group_by(word) %>%
  tally %>%
  arrange(desc(n)) %>%
  head(20) %>%
  mutate(`# papers` = countPapersUsingWord(word)) %>%
  add_column(place = c(1:nrow(.)), .before = 0)

top_words_2019
```

_What are the top word stems?_

```{r top_stems_2019}
countPapersUsingStem <- function(the_stem) {
  sapply(the_stem, function(s) {
    no_stop_stems_2019 %>%
      filter(word_stem == s) %>%
      group_by(id) %>%
      count %>%
      nrow
  })
}

top_stems_2019 <- no_stop_stems_2019 %>%
  group_by(word_stem) %>%
  tally %>%
  arrange(desc(n)) %>%
  head(20) %>%
  mutate(`# papers` = countPapersUsingStem(word_stem)) %>%
  add_column(place = c(1:nrow(.)), .before = 0)

top_stems_2019
```

### Word cloud based on word stems

```{r plot_2019,dpi=600,fig.width=7,fig.asp=0.85}
# minimum occurence manually tested so that all words could be plotted
wordStemPlot(no_stop_stems_2019, top_stems_2019, "2019", 145, full_papers_2019, other_papers_2019)
```

## Reproducibility keywords per year - absolute

```{r keywords_per_year_abs}
keywords_2018 <- word_counts_sums_total_2018
keywords_2019 <- word_counts_sums_total_2019

names(keywords_2018)[[1]] <- names(keywords_2019)[[1]] <- "year"
keywords_2018$year <- "2018"
keywords_2019$year <- "2019"

rbind(keywords_2018, keywords_2019)
```

## Reproducibility keywords per year - per paper

```{r keywords_per_year_per_paper}
cbind(year = c("2018", "2019"),
      round(rbind(
        dplyr::bind_cols(keywords_2018[-(1)] / (full_papers_2018 + other_papers_2018)),
        dplyr::bind_cols(keywords_2019[-(1)] / (full_papers_2019 + other_papers_2019))
        ), digits = 2)
)
```

## Reproducibility keywords per year - per 1000 words

```{r keywords_per_year_per_word}
cbind(year = c("2018", "2019"),
      round(rbind(
        dplyr::bind_cols(keywords_2018[-(1)] / nrow(no_stop_words_2018) * 1000),
        dplyr::bind_cols(keywords_2019[-(1)] / nrow(no_stop_words_2019) * 1000)
        ), digits = 2)
)
```

--------

## License

This document is licensed under a [Creative Commons Attribution 4.0 International License](https://creativecommons.org/licenses/by/4.0/).

All contained code is licensed under the [Apache License 2.0](https://choosealicense.com/licenses/apache-2.0/).

## Metadata

```{r session_info}
devtools::session_info(include_base = TRUE)
```
